---
title: "allfish_conditional"
author: "Sarah Roberts"
date: "2/2/2022"
output: html_document
---

First, predict out of sample by fitting the model with a training set and then predicting (conditional and traditional) the test set. 

Then predict in sample by fitting the model with the entire dataset and predicting (conditional and traditional) the data used to fit the model.

Compare conditional and traditional prediction for both in and out of sample, and plot.


#packages
```{r}
library(gjam)
library(tidyverse)
library(ggthemes)
library(ggpubr)
library(ggplot2)
library(pROC)
library(gridExtra)
library(sf)
```
#functions 
```{r}
r2_general <-function(preds,actual){ 
  return(1- sum((preds - actual) ^ 2)/sum((actual - mean(actual))^2))
}

RMSE_func <- function(preds, actual){
  return(sqrt(mean((actual - preds)^2)))
}

theme_Publication <- function(base_size=12, base_family="Arial") {

      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.5, "cm"),
               #legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}
'%!in%' <- function(x,y)!('%in%'(x,y))

.cleanNames <- function(xx){
  
  xx <- gsub('-','',xx)
  xx <- gsub('_','',xx)
  xx <- gsub(' ','',xx)
  xx <- gsub("'",'',xx)
  
  xx
}

conditionalComparison <- function( output, newdata ){
  
  # compare conditional prediction with unconditional prediction in output
  
  ydataCond <- newdata$ydataCond
  colnames(ydataCond) <- .cleanNames( colnames(ydataCond) )
  
  y           <- as.matrix( output$inputs$y )
  colnames(y) <- .cleanNames( colnames(y) )
  
  typeNames <- output$modelList$typeNames
  zeroCols  <- which(typeNames %in% c('DA','CA','OC','CC'))
  
  x         <- output$inputs$xUnstand
  S         <- ncol(y)
  n         <- nrow(x)
  if( is.null(ydataCond) )stop(' newdata must include ydataCond ' )
  if( is.null(colnames(ydataCond)) )stop( 'ydataCond must have colnames matching ydata' )
  
  y0 <- y
  y0[ y0 > 0 ] <- 1
  percZero <- signif(1 - apply(y0, 2, sum)/nrow(y0),3)*100
  
  ni <- match(colnames(ydataCond), colnames(y))   # condition on these columns
  ci <- c(1:S)[-ni]
  if( length(ni) < ncol(ydataCond) )stop( 'ydataCond must have colnames matching ydata' )
  
  edata <- output$inputs$effMat
  if( is.null(edata) )edata <- newdata$effort$values
  if( is.null(edata) )edata <- y*0 + 1
  wUn <- output$prediction$ypredMu/edata
  w   <- y/edata
  
  preds  <- gjamPredict(output, newdata = newdata, FULL=T) 
  condy  <- preds$sdList$yMu                                # on data scale
  condw  <- condy/edata                                     # latent scale
  
  coni <- (w[,ci] - condw[,ci] )^2
  unci <- ( w[,ci] - wUn[,ci] )^2
  umc  <- unci - coni
  uno  <- length(which(umc > 0))/length(umc)  # fraction larger error in uncon
  
  uno <- umc
  uno[ uno > 0 ] <- 1
  uno[ uno <= 0 ] <- 0
  
  
  if( length(ci) == 1 ){
    con <- mean( coni  )
    unc <- mean( unci )
    uno <- sum(uno)/length(uno)
  }else{ 
    con <- diag( var( w[,ci] - condw[,ci] ) )
    unc <- diag( var( w[,ci] - wUn[,ci] ) )
    uno <- apply(uno, 2, sum)/nrow(umc)
  }
  
  wi <- which(typeNames[ci] == 'PA')
  if( length(wi) > 0){
    tt <- .binaryScore(condw[,ci[wi]], y[,ci[wi]])
    con <- tt$brierScore
    #  ccLog   <- tt$logScore
    
    tt <- .binaryScore(wUn[,ci[wi]], y[,ci[wi]])
    unc <- tt$brierScore
    #  uuLog   <- tt$logScore
  }
  
  
  sigma <- output$parameters$sigMu
  beta  <- output$parameters$betaMu
  
  C  <- solve(sigma[ni,ni])%*%sigma[ni,ci]      # check for correlation matrix
  mu <- x%*%beta
  
  if(length(zeroCols) > 0){
    mz <- mu[,zeroCols]
    mz[ mz < 0 ] <- 0
    mu[,zeroCols] <- mz
  }
  Rs <- w[,ci,drop=F] - mu[,ci,drop=F]
  Rn <- w[,ni,drop=F] - mu[,ni,drop=F]
  D  <- Rn%*%C
  # z  <- -(2*t(Rs) + t(D))%*%D
  
  z <- ( 2*t(Rs) - t(C)%*%t(Rn) )%*%Rn%*%C
  colnames(z) <- rownames(z)
  
  vmu <- var(mu[,ci])
  vd  <- var(D)
  cf  <- vd/(vmu + vd)      #variance fraction from conditional
  
  fe <- output$fit$fractionExplained
  ynames <- colnames(fe)
  ynames <- ynames[ ynames %in% colnames(condy)[ci] ]
  ynames <- ynames[ !ynames == 'other' ]
  
  if( length(wi) == 0 ){   # not PA data
    
    if( length(ci) == 1 ){
      
      imat <- cbind( unc, con, 
                     (unc - con)/unc*100, uno,
                     fe[1,ynames], diag(cf), diag(z)[ynames]/n )
    }else{
      imat <- cbind( unc[ynames], con[ynames], 
                     (unc - con)[ynames]/unc[ynames]*100, uno[ynames],
                     fe[1,ynames], diag(cf)[ynames], diag(z)[ynames]/n )
    }
    colnames(imat) <- c("Unc MSPE", "Con MSPE", "Perc Diff", "Frac u > c",
                        "Frac Expl", 
                        "Frac Con", "CIS")
  }else{  # PA data
    
    if( length(ci) == 1 ){
      
      imat <- cbind( unc, con, 
                     (unc - con)/unc*100, uno,
                     fe[1,ynames], diag(cf) )
    }else{
      imat <- cbind( unc[ynames], con[ynames], (unc - con)[ynames]/unc[ynames]*100, 
                     uno[ynames],
                     fe[1,ynames], diag(cf)[ynames] )
    }
    colnames(imat) <- c("Unc Brier", "Con Brier", "Perc Diff", "Frac u > c",
                        "Frac Expl", 
                        "Frac Con")
  }
  if(typeNames[1] != 'CON'){
    percZero <- percZero[ynames]
    imat <- cbind(imat, percZero)
  }
  if( nrow(imat) == 1)rownames(imat) <- colnames(y)[ci]
  out1 <- list(imat, C)
  out1
}


```


Set up some basic parameters
```{r}
ng <- 20000
burnin <- ng/2

nsim <- 1000
```



# 1 out of sample 
train on 70% and test on 30% 

## 1.1 Fit model
After you fit the model, save it (and the train and test data) so you don't have to run it again.
```{r}

# load data
load("../DATA/fish/ydata.rdata")
load("../DATA/fish/xdata.rdata")


# get sqrt y
ydata <- sqrt(ydata)


# keep only data from Gulf of Maine
keep <- grep("GulfofMaine", xdata$SUBREGION2)
xdata <- xdata[keep,]
ydata <- ydata[keep,]

# keep only species with enough data
ydata <- ydata[,colSums(ydata != 0) > 150]

ynames <- colnames(ydata)
xnames <- colnames(xdata)

# # combine x and y, keep only observations from Gulf of Maine
total <- cbind(ydata, xdata)

# take out 70% for training set
smp_size <- floor(0.70 * nrow(total))
set.seed(123)
train_ind <- sample(seq_len(nrow(total)), size = smp_size)
train <- total[train_ind, ]
test <- total[-train_ind, ]

ytrain <- train[,colnames(train) %in% ynames]
xtrain <- train[,colnames(train) %in% xnames]
ytest <- test [,colnames(test ) %in% ynames]
xtest <- test [,colnames(test) %in% xnames]

#run gjam

ml <- list(ng = ng, burnin = burnin, typeNames = 'CA')
out_oos <- gjam(~BT + depth + BSAL + SST + chla + SEDSIZE + month, xdata = xtrain, ydata = ytrain, modelList = ml) 

save(out_oos, ytrain, ytest, xtrain, xtest, file='../OUT/fish-out_train_CPUE_fall.rdata')


```


## 1.2 Conditional prediction
Conditional predict test data. Each species is conditioned on all other species
```{r}

load("../OUT/fish-out_train_CPUE_fall.rdata")

#conditionally predict
#holdout
row_names <- rownames(ytest)
col_names <- colnames(ytest)
pred_con <- matrix(nrow = nrow(ytest), ncol = ncol(ytest), dimnames = list(row_names, col_names))

pb <- txtProgressBar(0, ncol(ytest), style = 3)

for(p in 1:ncol(ytest)){
  pdex <- colnames(ytest)[p]
  
  # conditionally predict out-of-sample
  newdata <- list(xdata = xtest, ydataCond = ytest[,-p], nsim = nsim) 

  p6      <- gjamPredict(out_oos, newdata = newdata)
  pred_con[,pdex] <- p6$sdList$yMu[,p]
  
  setTxtProgressBar(pb, p)
  Sys.sleep(time = 1)
}

```


## 1.3 Traditional prediction
Traditional predict test data.
```{r}

#traditional predict
newdata <- list(xdata = xtest, nsim=nsim)
p7      <- gjamPredict(out_oos, newdata = newdata)
y_obs <- ytest
y_pred <- p7[["sdList"]][["yMu"]]
```


## 1.4 ConditionalComparison

Compare out of sample conditional and traditional. 
This section modifies the function `conditionalComparison()` for use with out-of-sample prediction
```{r}

y <- ytest

# predictions made in previous sections
y_con <- pred_con
y_uncon <- y_pred

# compare conditional prediction with unconditional prediction in output

ydataCond <- y_con
colnames(ydataCond) <- .cleanNames( colnames(ydataCond) )

colnames(y) <- .cleanNames( colnames(y) )

# typeNames <- output$modelList$typeNames
# zeroCols  <- which(typeNames %in% c('DA','CA','OC','CC'))

x         <- xtest
S         <- ncol(y)
n         <- nrow(x)

y0 <- y
y0[ y0 > 0 ] <- 1
percZero <- signif(1 - apply(y0, 2, sum)/nrow(y0),3)*100

ni <- match(colnames(ydataCond), colnames(y))   # condition on these columns
ci <- c(1:S)[-ni]
ci <- c(1:S)
#if( length(ni) < ncol(ydataCond) )stop( 'ydataCond must have colnames matching ydata' )

edata <- y*0 + 1
wUn <- y_uncon#/edata
w   <- y#/edata

#preds  <- gjamPredict(output, newdata = newdata, FULL=T) 
condy  <- y_con                                # on data scale
condw  <- condy#/edata                                     # latent scale

coni <- (w[,ci] - condw[,ci] )^2
unci <- ( w[,ci] - wUn[,ci] )^2
umc  <- unci - coni
uno  <- length(which(umc > 0))/length(umc)  # fraction larger error in uncon

uno <- umc
uno[ uno > 0 ] <- 1
uno[ uno <= 0 ] <- 0


if( length(ci) == 1 ){
  con <- mean( coni  )
  unc <- mean( unci )
  uno <- sum(uno)/length(uno)
}else{ 
  con <- diag( var( w[,ci] - condw[,ci] ) )
  unc <- diag( var( w[,ci] - wUn[,ci] ) )
  uno <- apply(uno, 2, sum)/nrow(umc)
}


imat <- cbind( unc, con, 
               (unc - con)/unc*100, uno)
colnames(imat) <- c("Unc MSPE", "Con MSPE", "Perc Diff", "Frac u > c")

```


# 2 in sample

## 2.1 Fit model

Combine test and train data and fit model.
```{r}

# load data
load("../DATA/fish/ydata.rdata")
load("../DATA/fish/xdata.rdata")


# get sqrt y
ydata <- sqrt(ydata)


# keep only data from Gulf of Maine
keep <- grep("GulfofMaine", xdata$SUBREGION2)
xdata <- xdata[keep,]
ydata <- ydata[keep,]

# keep only species with enough data
ydata <- ydata[,colSums(ydata != 0) > 150]

ynames <- colnames(ydata)
xnames <- colnames(xdata)

# # combine x and y, keep only observations from Gulf of Maine
total <- cbind(ydata, xdata)


#run gjam
ml <- list(ng = ng, burnin = burnin, typeNames = 'CA')
out_is <- gjam(~BT + depth + BSAL + SST + chla + SEDSIZE + month, xdata = xdata, ydata = ydata, modelList = ml) 

save(out_is, xdata, ydata, file='../OUT/fish-out_full_CPUE_fall.Rdata')

```



## 2.2 ConditionalComparison

Condition each species on all others
```{r}
load('../ouT/fish-out_full_CPUE_fall.Rdata') #in case you don't want to run the full thing


all <- c() # store comparison metrics
pred_con <- c() # store conditional predictions
out <- out_is

ydata <- out$inputs$y
sp <- colnames(ydata)

for (s in 1:length(sp)) {
  print(paste0("Species ", s, " of ", length(sp)))
  
  pred <- sp[s]
  condOn <- sp[-s]
  
  condOnData <- ydata[,condOn]
  newdata <- list(ydataCond = condOnData,  nsim=nsim)
  
  # compare conditional and unconditional
  comp <- conditionalComparison( out, newdata )
  comp <- as.data.frame(comp[[1]])
  
  comp$species <- sp[s]
  
  all <- bind_rows(all, comp)
  
  
  # predict each species conditionally
  tmp <- gjamPredict(out, newdata = newdata)
  pred_con <- bind_cols(pred_con, tmp$sdList$yMu[,pred])
  colnames(pred_con)[s] <- pred
}



```


#3 plot it 
plot conditional and unconditional metrics together
```{r}

# format in sample metrics
all1 <- all %>%
  group_by(species) %>%
  summarize(percDiffMean = mean(`Perc Diff`),
            percDiffSe = sd(`Perc Diff`)/sqrt(n()),
            percImprMean = mean(`Frac u > c`*100),
            percImprSe = sd(`Frac u > c`*100)/sqrt(n())) %>%
  pivot_longer(cols = c("percDiffMean", "percImprMean", 
                        "percDiffSe", "percImprSe")) %>%
  mutate(type = case_when(name %in% c("percDiffMean", "percImprMean") ~ "mean",
                          T ~ "se"),
         metric = case_when(name %in% c("percDiffMean", "percDiffSe") ~ "percDiff",
                            T ~ "percImpr")) %>%
  select(!name) %>%
  pivot_wider(names_from = type, values_from = value) 


# format out of sample metrics
imat <- as.data.frame(imat)
imat$species <- rownames(imat)
imat$`Frac u > c` <- imat$`Frac u > c`*100

imat1 <- imat[,c("Perc Diff", "Frac u > c", "species")]
colnames(imat1) <- c("percDiff", "percImpr", "species")
imat1 <- imat1 %>%
  pivot_longer(!species)
colnames(imat1) <- c("species", "metric", "oos")

colnames(all1) <- c("species", "metric", "is")
all1 <- all1[,1:3]

# calculate number of zeros
yzeros <- ydata
yzeros[yzeros > 0] <- 1
zeros <- as.data.frame(colMeans(yzeros))
zeros$species <- rownames(zeros)
colnames(zeros) <- c("abun", "species")

# join everything together
full <- full_join(all1, imat1, by = c("species", "metric")) %>%
  full_join(zeros, "species") %>%
  pivot_longer(cols = c("is", "oos")) %>%
  mutate(Species = case_when(species == "ACADIAN.REDFISH" ~ "Acadian Redfish",
                             species == "AMERICAN.LOBSTER" ~ "American Lobster",
                             species == "AMERICAN.PLAICE" ~ "American Plaice",
                             species == "ATLANTIC.COD" ~ "Atlantic Cod",
                             species == "ATLANTIC.HERRING" ~ "Atlantic Herring",
                             species == "BARNDOOR.SKATE" ~ "Barndoor Skate",
                             species == "BUTTERFISH" ~ "Butterfish",
                             species == "FOURSPOT.FLOUNDER" ~ "Four-spot Flounder",
                             species == "HADDOCK" ~ "Haddock",
                             species == "LITTLE.SKATE" ~ "Little Skate",
                             species == "LONGFIN.SQUID" ~ "Longfin Squid",
                             species == "LONGHORN.SCULPIN" ~ "Longhorn Sculpin",
                             species == "NORTHERN.SEAROBIN" ~ "Northern Searobin",
                             species == "NORTHERN.SHORTFIN.SQUID" ~ "Northern Shortfin Squid",
                             species == "POLLOCK" ~ "Pollock",
                             species == "RED.HAKE" ~ "Red Hake",
                             species == "SEA.SCALLOP" ~ "Sea Scallop",
                             species == "SILVER.HAKE" ~ "Silver Hake",
                             species == "SPINY.DOGFISH" ~ "Spinny Dogfish",
                             species == "SUMMER.FLOUNDER" ~ "Summer Flounder",
                             species == "WHITE.HAKE" ~ "White Hake",
                             species == "WINTER.FLOUNDER" ~ "Winter Flounder",
                             species == "WINTER.SKATE" ~ "Winter Skate",
                             species == "YELLOWTAIL.FLOUNDER" ~ "Yellowtail Flounder"),
         group = case_when(name == "is" ~ "In-sample",
                           name == "oos" ~ "Out-of-sample")) 


b <- ggplot(full) + 
  geom_hline(yintercept = 50, color = "deepskyblue", alpha = 0.5) +
  geom_hline(yintercept = 0, color = "coral3", alpha = 0.5) +
  geom_point(aes(x = reorder(Species, abun), y = value, color = metric, shape = name)) +
  scale_color_manual(#labels = c("% Improvement in RMSPE \nfrom Conditioning",
    #          "Fraction of Predictions\nImproved by Conditioning"),
    values = c("coral3","deepskyblue"), guide = "none") +
  scale_shape_manual(values = c(16, 1), guide = "none") +
  theme_bw() +
  theme(strip.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(1.5, "lines"),
        plot.margin = unit(c(1,.1,0,0), "cm"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1,
                                   size = 7),
        axis.text.y = element_blank(),
        axis.title = element_text(size = 8)) +
  coord_cartesian(ylim = c(0, 100), clip = "on") +
  labs(x = "", y = "")

# b <- ggplot(filter(full, metric == "percImpr")) +
#   geom_boxplot(aes(x = name, y = value, alpha = name),
#                color = "deepskyblue",
#                fill = "deepskyblue") +
#   scale_alpha_manual(values = c(.6, 0)) +
#   scale_y_continuous(breaks = c(50, 75, 100)) +
#   theme_bw() +
#   theme(strip.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.spacing = unit(1.5, "lines"),
#         plot.margin = margin(1, 0, 0, 0, "cm"),
#         axis.text.x = element_blank(),
#         axis.ticks.x = element_blank(),
#         legend.position = "none") +
#   coord_cartesian(ylim = c(50, 100), clip = "on") +
#   labs(x = "", y = "")
# 
# c <- ggplot(filter(full, metric == "percDiff")) +
#   geom_boxplot(aes(x = name, y = value, alpha = name),
#                color = "coral3",
#                fill = "coral3") +
#   scale_alpha_manual(values = c(.6, 0)) +
#   scale_y_continuous(breaks = c(0, 25, 50)) +
#   theme_bw() +
#   theme(strip.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.spacing = unit(1.5, "lines"),
#         plot.margin = margin(0, 0, 1, 0, "cm"),
#         axis.ticks.x = element_blank(),
#         legend.position = "none") +
#   coord_cartesian(ylim = c(0, 50), clip = "on") +
#   labs(x = "", y = "")
# 


a <- ggplot() + 
  geom_hline(yintercept = 50, color = "deepskyblue", alpha = 0.5) +
  geom_hline(yintercept = 0, color = "coral3", alpha = 0.5) +
  geom_boxplot(data = filter(full, metric == "percDiff"),
               aes(x = group, y = value, alpha = name),
               color = "coral3",
               fill = "coral3") +
  geom_boxplot(data = filter(full, metric == "percImpr"),
               aes(x = group, y = value, alpha = name),
               fill = "deepskyblue",
               color = "deepskyblue") +
  scale_alpha_manual(values = c(.6, 0)) +
  coord_cartesian(ylim = c(0, 100)) +
  theme_bw() +
  theme(strip.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(1.5, "lines"),
        plot.margin = unit(c(0, 0, 1, 0), "cm"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1,
                                   size = 7),
        axis.text.y = element_text(size = 7),
        axis.title = element_text(size = 8),
        legend.position = "none") +
  labs(x = "", y = "")
  



ggpubr::ggarrange(a, b, nrow = 1,
                  labels = "auto",
                  widths = c(0.3, 1),
                  label.x = c(0.04, 0.02), label.y = .95,
                  font.label = list(size = 10),
                  align = "hv")


# library(patchwork)
# (b/c) + a +
#   plot_layout(widths = c(1, 0.3))

ggsave(filename = "../OUT/figures/metrics.png",
       height = 100, width = 180, units = "mm", dpi = 600)

```

#4 Map

```{r}
## Map ----

# map in sample
output <- out_is
ydata <- output$inputs$y
xdata <- output$inputs$xdata



# format xdata with lat and longs
total1 <- read.csv("../DATA/fish/total_with_env.csv") #this has the missing lat longs in it 

colnames(total1) <- gsub("\\_", ".", colnames(total1)) #gjam doesnt like _ in column names 
total1 <- total1[total1$year>1986,] 
total1 <- total1[total1$SEASON == "FALL",]

total1$SEDSIZE <- ifelse(total1$SEDIMENT ==  '0 - 0.03 Silt/Mud', .03, 
                        ifelse(total1$SEDIMENT ==  '0.03 - 0.17 Sand', .17,
                               ifelse(total1$SEDIMENT ==  '0.17 - 0.35 Sand',.35, 
                                      ifelse(total1$SEDIMENT ==  '0.35 - 0.36 Sand', .36,
                                             ifelse(total1$SEDIMENT ==  '0.36 - 0.48 Sand', .48, ifelse(total1$SEDIMENT ==  '0.48+ Coarse Sand to Gravel',.50,NA))))))

xnames <- c("depth", "BT", "BSAL", "SSAL", "SST", "chla", "month", "season", "SEDSIZE", "SUBREGION2", "haulid", "lat", "lon", "stratum", "stratumarea", "year")
xdata1 <- total1[colnames(total1) %in% xnames]
xdata1 <- xdata1[xdata1$SUBREGION2 == "Gulf of Maine",]
xdata1$SUBREGION2 <- gsub("Gulf of Maine", "GulfofMaine", xdata1$SUBREGION2 )

xdata1$season <- as.factor(xdata1$season)
xdata1$SUBREGION2 <- as.factor(xdata1$SUBREGION2)

xdata <- left_join(xdata, xdata1,by = c("depth", "BT", "BSAL", "SSAL", "SST", "chla", "month", "season", "SEDSIZE", "SUBREGION2"))



### get predictions together ----

## observed
total_strat_obs <- cbind(xdata, ydata) 
total_strat_obs <- total_strat_obs %>% pivot_longer(cols = colnames(ydata), names_to = "spp", values_to = "biomass")

total_strat_obs$pred_typ <- "observed"

## conditional prediction
total_strat_con <- cbind(xdata, pred_con) 
total_strat_con <- total_strat_con %>% pivot_longer(cols = colnames(pred_con), names_to = "spp", values_to = "biomass")


total_strat_con$pred_typ <- "conditional"

## unconditional prediction
pred_uncon <- out$prediction$ypredMu
total_strat_uncon <- cbind(xdata, pred_uncon) 
total_strat_uncon <- total_strat_uncon %>% pivot_longer(cols = colnames(pred_uncon), names_to = "spp", values_to = "biomass")


total_strat_uncon$pred_typ <- "unconditional"

total_strat <- rbind(total_strat_obs, total_strat_con, total_strat_uncon)


## combine
total_strat <- total_strat %>% 
  group_by(season, spp, year, stratum, stratumarea, pred_typ) %>% 
  summarise(biomass = mean(biomass)) %>% 
  mutate(biomass_area = biomass*stratumarea,
         pred_type = factor(pred_typ, levels = c("observed", "unconditional", "conditional")))


### plot ----
strat <- sf::st_read("../DATA/fish/stratum/BTS_Strata.shp")
#plot(strat)

strat <- left_join(strat, total_strat, by = c("STRATA" = "stratum"))
st_crs(strat) = 4326 #this is WGS 1984


### plot ----
library(rnaturalearth)
world <- ne_countries(scale = "medium", returnclass = "sf")

#### map of predicted abundances
# cols <- c("blue3", 
#           "purple",
#           "orange", "brown")
# p <- ggplot() + 
#   geom_sf(data = world) + 
#   geom_sf(data = subset(strat, 
#                         strat$spp == "ATLANTIC.COD" & strat$year == 2016),
#           aes(group = as.factor(biomass_area), fill = biomass_area)) + 
#   coord_sf(xlim=c(-74, -65), ylim=c(40,46), expand = TRUE) +
#   scale_fill_gradientn(colours = cols,
#                         values = c(0, 0.1, 0.35, 0.75, 1)) + 
#   facet_wrap(~ pred_type) + 
#   theme_Publication() + 
#   theme(legend.position = "right", legend.direction = "vertical") +
#   labs(fill = "Atlantic cod\nbiomass per stratum\n(kg/km2)")+
#   theme_bw() +
#   theme(panel.grid = element_blank(),
#         strip.background = element_blank(),
#         strip.text = element_text(hjust = 0),
#         axis.title = element_blank(),
#         panel.spacing.x = unit(6, "mm")) 
# 
# ggsave("codmaps_1.jpg", plot=p, width = 25, height=10, units=c("cm"), dpi=500)

#### difference maps 
strat_sub <- strat %>% 
  dplyr::select(-pred_typ, -biomass) %>% 
  pivot_wider(names_from = "pred_type", values_from = "biomass_area") %>% 
  mutate(sub_con = conditional - observed, 
         sub_uncon = unconditional - observed) %>%
  pivot_longer(cols = c("sub_uncon", "sub_con", "observed"))

strat_sub <- st_as_sf(strat_sub)
strat_sub <- strat_sub %>% filter(name %in% c("sub_uncon", "sub_con", "observed"))
st_crs(strat_sub) = 4326
strat_sub$name1 <- as.factor(strat_sub$name)


# just for the maps, set limits at 2000 and -1000
strat_sub$value[which(strat_sub$value > 2000)] <- 2000
strat_sub$value[which(strat_sub$value < -1000)] <- -1000



base <- ggplot() + 
  geom_sf(data = world) + 
  labs(fill = "Atlantic cod biomass (kg/km2)")+
  theme_bw() +
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(hjust = 0),
        plot.title = element_text(size = 8,
                             hjust = 0.5),
        axis.title = element_blank(),
        axis.text = element_text(size = 7),
        #panel.spacing.x = unit(6, "mm"),
        plot.margin = margin(5, 1, 3, 1),
        legend.key.height = unit(.5, "cm"),
        legend.key.width = unit(1.1, "cm"),
        legend.title = element_text(size = 7,
                                    hjust = 0.5),
        #legend.title.align = 0.5,
        legend.direction = "horizontal",
        legend.text= element_text(size = 6)) +
  scale_fill_gradient2(low = "darkblue", high = "red4", mid = "white", 
                      midpoint = 0, breaks = c(-1000,0,1000,2000),
                      limits = c(-1000, 2000),
                      guide = guide_colorbar(title.position = "top")) +
  scale_y_continuous(breaks = c(40, 43, 46)) +
  scale_x_continuous(breaks = c(-66, -70, -74))



a <- base +
  geom_sf(data = filter(strat_sub, 
                        strat_sub$spp == "ATLANTIC.COD",
                        strat_sub$year == 2012,
                        name == "observed"),
          aes(group = as.factor(value), fill = value)) +
  coord_sf(xlim=c(-74, -65), ylim=c(40,46), expand = TRUE) +
  labs(title = "Observed")
  
b <- base +
  geom_sf(data = filter(strat_sub, 
                        strat_sub$spp == "ATLANTIC.COD",
                        strat_sub$year == 2012,
                        name == "sub_uncon"),
          aes(group = as.factor(value), fill = value)) +
  coord_sf(xlim=c(-74, -65), ylim=c(40,46), expand = TRUE) +
  labs(title = "Traditional Prediction - Observed")
  
c <- base +
  geom_sf(data = filter(strat_sub, 
                        strat_sub$spp == "ATLANTIC.COD",
                        strat_sub$year == 2012,
                        name == "sub_con"),
          aes(group = as.factor(value), fill = value)) +
  coord_sf(xlim=c(-74, -65), ylim=c(40,46), expand = TRUE) +
  labs(title = "Conditional Prediction - Observed")
  
ggpubr::ggarrange(a, b, c, ncol = 1,
                  labels = "auto", common.legend = T,
                  label.y = 1.01, label.x = 0.15,
                  legend = "bottom",
                  font.label = list(size = 10))

ggsave("../OUT/figures/codmaps.png", width = 80, height=180, units=c("mm"), 
       dpi = 600)
```
