---
title: "allfish_conditional"
author: "Sarah Roberts"
date: "2/2/2022"
output: html_document
---

holdout one species at a time and conditionally predict based on env. and all other species.

given xdata_test, xdata_train, ydata_test, ydata_train and the models (out_train_CPUEfall and out_full_CPUEfall)

When I updated my R and my mac to mojave, gjam predictions were really bad for all of the below examples. Before I updated, predictions were not all 0. 

#packages
```{r}
library(gjam)
library(tidyverse)
library(ggthemes)
library(ggpubr)
library(ggplot2)
library(pROC)
library(gridExtra)
library(sf)
```
#functions 
```{r}
r2_general <-function(preds,actual){ 
  return(1- sum((preds - actual) ^ 2)/sum((actual - mean(actual))^2))
}

RMSE_func <- function(preds, actual){
  return(sqrt(mean((actual - preds)^2)))
}

theme_Publication <- function(base_size=12, base_family="Arial") {

      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.5, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}
'%!in%' <- function(x,y)!('%in%'(x,y))

.cleanNames <- function(xx){
  
  xx <- gsub('-','',xx)
  xx <- gsub('_','',xx)
  xx <- gsub(' ','',xx)
  xx <- gsub("'",'',xx)
  
  xx
}

conditionalComparison <- function( output, newdata ){
  
  # compare conditional prediction with unconditional prediction in output
  
  ydataCond <- newdata$ydataCond
  colnames(ydataCond) <- .cleanNames( colnames(ydataCond) )
  
  y           <- as.matrix( output$inputs$y )
  colnames(y) <- .cleanNames( colnames(y) )
  
  typeNames <- output$modelList$typeNames
  zeroCols  <- which(typeNames %in% c('DA','CA','OC','CC'))
  
  x         <- output$inputs$xUnstand
  S         <- ncol(y)
  n         <- nrow(x)
  if( is.null(ydataCond) )stop(' newdata must include ydataCond ' )
  if( is.null(colnames(ydataCond)) )stop( 'ydataCond must have colnames matching ydata' )
  
  y0 <- y
  y0[ y0 > 0 ] <- 1
  percZero <- signif(1 - apply(y0, 2, sum)/nrow(y0),3)*100
  
  ni <- match(colnames(ydataCond), colnames(y))   # condition on these columns
  ci <- c(1:S)[-ni]
  if( length(ni) < ncol(ydataCond) )stop( 'ydataCond must have colnames matching ydata' )
  
  edata <- output$inputs$effMat
  if( is.null(edata) )edata <- newdata$effort$values
  if( is.null(edata) )edata <- y*0 + 1
  wUn <- output$prediction$ypredMu/edata
  w   <- y/edata
  
  preds  <- gjamPredict(output, newdata = newdata, FULL=T) 
  condy  <- preds$sdList$yMu                                # on data scale
  condw  <- condy/edata                                     # latent scale
  
  coni <- (w[,ci] - condw[,ci] )^2
  unci <- ( w[,ci] - wUn[,ci] )^2
  umc  <- unci - coni
  uno  <- length(which(umc > 0))/length(umc)  # fraction larger error in uncon
  
  uno <- umc
  uno[ uno > 0 ] <- 1
  uno[ uno <= 0 ] <- 0
  
  
  if( length(ci) == 1 ){
    con <- mean( coni  )
    unc <- mean( unci )
    uno <- sum(uno)/length(uno)
  }else{ 
    con <- diag( var( w[,ci] - condw[,ci] ) )
    unc <- diag( var( w[,ci] - wUn[,ci] ) )
    uno <- apply(uno, 2, sum)/nrow(umc)
  }
  
  wi <- which(typeNames[ci] == 'PA')
  if( length(wi) > 0){
    tt <- .binaryScore(condw[,ci[wi]], y[,ci[wi]])
    con <- tt$brierScore
    #  ccLog   <- tt$logScore
    
    tt <- .binaryScore(wUn[,ci[wi]], y[,ci[wi]])
    unc <- tt$brierScore
    #  uuLog   <- tt$logScore
  }
  
  
  sigma <- output$parameters$sigMu
  beta  <- output$parameters$betaMu
  
  C  <- solve(sigma[ni,ni])%*%sigma[ni,ci]      # check for correlation matrix
  mu <- x%*%beta
  
  if(length(zeroCols) > 0){
    mz <- mu[,zeroCols]
    mz[ mz < 0 ] <- 0
    mu[,zeroCols] <- mz
  }
  Rs <- w[,ci,drop=F] - mu[,ci,drop=F]
  Rn <- w[,ni,drop=F] - mu[,ni,drop=F]
  D  <- Rn%*%C
  # z  <- -(2*t(Rs) + t(D))%*%D
  
  z <- ( 2*t(Rs) - t(C)%*%t(Rn) )%*%Rn%*%C
  colnames(z) <- rownames(z)
  
  vmu <- var(mu[,ci])
  vd  <- var(D)
  cf  <- vd/(vmu + vd)      #variance fraction from conditional
  
  fe <- output$fit$fractionExplained
  ynames <- colnames(fe)
  ynames <- ynames[ ynames %in% colnames(condy)[ci] ]
  ynames <- ynames[ !ynames == 'other' ]
  
  if( length(wi) == 0 ){   # not PA data
    
    if( length(ci) == 1 ){
      
      imat <- cbind( unc, con, 
                     (unc - con)/unc*100, uno,
                     fe[1,ynames], diag(cf), diag(z)[ynames]/n )
    }else{
      imat <- cbind( unc[ynames], con[ynames], 
                     (unc - con)[ynames]/unc[ynames]*100, uno[ynames],
                     fe[1,ynames], diag(cf)[ynames], diag(z)[ynames]/n )
    }
    colnames(imat) <- c("Unc MSPE", "Con MSPE", "Perc Diff", "Frac u > c",
                        "Frac Expl", 
                        "Frac Con", "CIS")
  }else{  # PA data
    
    if( length(ci) == 1 ){
      
      imat <- cbind( unc, con, 
                     (unc - con)/unc*100, uno,
                     fe[1,ynames], diag(cf) )
    }else{
      imat <- cbind( unc[ynames], con[ynames], (unc - con)[ynames]/unc[ynames]*100, 
                     uno[ynames],
                     fe[1,ynames], diag(cf)[ynames] )
    }
    colnames(imat) <- c("Unc Brier", "Con Brier", "Perc Diff", "Frac u > c",
                        "Frac Expl", 
                        "Frac Con")
  }
  if(typeNames[1] != 'CON'){
    percZero <- percZero[ynames]
    imat <- cbind(imat, percZero)
  }
  if( nrow(imat) == 1)rownames(imat) <- colnames(y)[ci]
  out1 <- list(imat, C)
  out1
}


```

# CPUE (counts per unit effort)

# 1 out of sample 
train on 70% and test on 30% 

## 1.1 Fit model
If you don't want to re-fit the model, the output is already saved and can be loaded in the next chunk.
```{r}

#read data 
ytest <- read.csv("X:/conditionalPrediction/DATA/fish/ytest.csv")
xtest <- read.csv("X:/conditionalPrediction/DATA/fish/xtest.csv")
ytrain <- read.csv("X:/conditionalPrediction/DATA/fish/ytrain.csv")
xtrain <- read.csv("X:/conditionalPrediction/DATA/fish/xtrain.csv")

ydata <- rbind(ytrain, ytest)
ydata <- sqrt(ydata)
xdata <- rbind(xtrain, xtest)

ynames <- colnames(ydata)
xnames <- colnames(xdata)


total <- cbind(ydata, xdata)
total <- total %>% dplyr::filter(SUBREGION2 == "GulfofMaine")

ydata <- total[,colnames(total) %in% ynames]
xdata <- total[,colnames(total) %in% xnames]
ydata <- ydata[,colSums(ydata != 0) > 150] # species with more than 150 observations

ynames <- colnames(ydata)
xnames <- colnames(xdata)

smp_size <- floor(0.70 * nrow(total))
set.seed(123)
train_ind <- sample(seq_len(nrow(total)), size = smp_size)
train <- total[train_ind, ]
test <- total[-train_ind, ]

ytrain <- train[,colnames(train) %in% ynames]
xtrain <- train[,colnames(train) %in% xnames]
ytest <- test [,colnames(test ) %in% ynames]
xtest <- test [,colnames(test) %in% xnames]

#run gjam

ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
out <- gjam(~BT + depth + BSAL + SST + chla + SEDSIZE + month, xdata = xtrain, ydata = ytrain, modelList = ml) 

save(out, file='X:/conditionalPrediction/OUT/fish-out_train_CPUE_fall.rdata')

```


## 1.2 Conditional prediction
Conditional predict test data. Each species is conditioned on all other species
```{r}
load('X:/conditionalPrediction/OUT/fish-out_train_CPUE_fall.Rdata') #in case you don't want to run the full thing
ytest <- read.csv("X:/conditionalPrediction/DATA/fish/ytest.csv")
xtest <- read.csv("X:/conditionalPrediction/DATA/fish/xtest.csv")

# only keep species that were fitted in the model
ytest <- ytest[,which(colnames(ytest) %in% colnames(out$inputs$y))]

#conditionally predict
#holdout
row_names <- rownames(ytest)
col_names <- colnames(ytest)
pred_con <- matrix(nrow = nrow(ytest), ncol = ncol(ytest), dimnames = list(row_names, col_names))

pb <- txtProgressBar(0, ncol(ytest), style = 3)

for(p in 1:ncol(ytest)){
  pdex <- colnames(ytest)[p]
  
  #ydataCond <- select(ytest, !all_of(pdex))
  
  newdata <- list(xdata = xtest, ydataCond = ytest[,-p], nsim = 300) # conditionally predict out-of-sample

  p6      <- gjamPredict(out, newdata = newdata)
  pred_con[,pdex] <- p6$sdList$yMu[,p]
  
  setTxtProgressBar(pb, p)
  Sys.sleep(time = 1)
}

```


## 1.3 Traditional prediction
Traditional predict test data.
```{r}

#traditional predict
newdata <- list(xdata = xtest, nsim=1000)
p7      <- gjamPredict(out, newdata = newdata)
y_obs <- ytest♥
y_pred <- p7[["sdList"]][["yMu"]]
```


## 1.4 ConditionalComparison

Compare out of sample conditional and traditional. 
This section modifies the function `conditionalComparison()` for use with out-of-sample prediction
```{r}

y_con <- pred_con
y_uncon <- p7[["sdList"]][["yMu"]]

# compare conditional prediction with unconditional prediction in output
  
  ydataCond <- y_con
  colnames(ydataCond) <- .cleanNames( colnames(ydataCond) )
  
  y           <- ytest
  colnames(y) <- .cleanNames( colnames(y) )
  
  # typeNames <- output$modelList$typeNames
  # zeroCols  <- which(typeNames %in% c('DA','CA','OC','CC'))
  
  x         <- xtest
  S         <- ncol(y)
  n         <- nrow(x)
  
  y0 <- y
  y0[ y0 > 0 ] <- 1
  percZero <- signif(1 - apply(y0, 2, sum)/nrow(y0),3)*100
  
  ni <- match(colnames(ydataCond), colnames(y))   # condition on these columns
  ci <- c(1:S)[-ni]
  ci <- c(1:S)
  #if( length(ni) < ncol(ydataCond) )stop( 'ydataCond must have colnames matching ydata' )
  
  edata <- y*0 + 1
  wUn <- y_uncon#/edata
  w   <- y#/edata
  
  #preds  <- gjamPredict(output, newdata = newdata, FULL=T) 
  condy  <- y_con                                # on data scale
  condw  <- condy#/edata                                     # latent scale
  
  coni <- (w[,ci] - condw[,ci] )^2
  unci <- ( w[,ci] - wUn[,ci] )^2
  umc  <- unci - coni
  uno  <- length(which(umc > 0))/length(umc)  # fraction larger error in uncon
  
  uno <- umc
  uno[ uno > 0 ] <- 1
  uno[ uno <= 0 ] <- 0
  
  
  if( length(ci) == 1 ){
    con <- mean( coni  )
    unc <- mean( unci )
    uno <- sum(uno)/length(uno)
  }else{ 
    con <- diag( var( w[,ci] - condw[,ci] ) )
    unc <- diag( var( w[,ci] - wUn[,ci] ) )
    uno <- apply(uno, 2, sum)/nrow(umc)
  }
  
  
      imat <- cbind( unc, con, 
                     (unc - con)/unc*100, uno)
    colnames(imat) <- c("Unc MSPE", "Con MSPE", "Perc Diff", "Frac u > c")

```


# 2 in sample

## 2.1 Fit model

Combine test and train data and fit model.
```{r}

#read data 
ytest <- read.csv("X:/conditionalPrediction/DATA/fish/ytest.csv")
xtest <- read.csv("X:/conditionalPrediction/DATA/fish/xtest.csv")
ytrain <- read.csv("X:/conditionalPrediction/DATA/fish/ytrain.csv")
xtrain <- read.csv("X:/conditionalPrediction/DATA/fish/xtrain.csv")

ydata <- rbind(ytrain, ytest)
ydata <- sqrt(ydata)
xdata <- rbind(xtrain, xtest)

ynames2 <- colnames(ydata)
xnames <- colnames(xdata)


total <- cbind(ydata, xdata)
total <- total %>% dplyr::filter(SUBREGION2 == "GulfofMaine")

ydata <- total[,colnames(total) %in% ynames]
xdata <- total[,colnames(total) %in% xnames]
ydata <- ydata[,colSums(ydata != 0) > 150] # I want non-zero rows 
ynames2 <- colnames(ydata)
#run gjam
ml <- list(ng = 20000, burnin = 8000, typeNames = 'CA')
out1 <- gjam(~BT + depth + BSAL + SST + chla + SEDSIZE + month, xdata = xdata, ydata = ydata, modelList = ml) 

save(out1, file='X:/conditionalPrediction/OUT/out_full_CPUE_fall.Rdata')

```



## 2.2 ConditionalComparison

```{r}
load('out_full_CPUE_fall.Rdata') #in case you don't want to run the full thing

all <- c()
  out <- out1

  ydata <- out$inputs$y
  sp <- colnames(ydata)
  
  for (s in 1:length(sp)) {
    print(paste0("Species ", s, " of ", length(sp)))
    
    pred <- sp[s]
    condOn <- sp[-s]
    
    condOnData <- ydata[,condOn]
    newdata <- list(ydataCond = condOnData,  nsim=200)
    
    # compare conditional and unconditional
    comp <- conditionalComparison( out, newdata )
    comp <- as.data.frame(comp[[1]])
    
    comp$species <- sp[s]
    
    all <- bind_rows(all, comp)
    
    
  }
  


```


#3 plot it 
plot conditional and unconditional metrics together
```{r}

all1 <- all %>%
  group_by(species) %>%
  summarize(percDiffMean = mean(`Perc Diff`),
            percDiffSe = sd(`Perc Diff`)/sqrt(n()),
            percImprMean = mean(`Frac u > c`*100),
            percImprSe = sd(`Frac u > c`*100)/sqrt(n())) %>%
  pivot_longer(cols = c("percDiffMean", "percImprMean", "percDiffSe", "percImprSe")) %>%
  mutate(type = case_when(name %in% c("percDiffMean", "percImprMean") ~ "mean",
                          T ~ "se"),
         metric = case_when(name %in% c("percDiffMean", "percDiffSe") ~ "percDiff",
                            T ~ "percImpr")) %>%
  select(!name) %>%
  pivot_wider(names_from = type, values_from = value) 


# 
# p1 <- ggplot(all1) +
#   geom_hline(yintercept = 50, color = "coral3", alpha = 0.5) +
#   geom_hline(yintercept = 0, color = "deepskyblue", alpha = 0.5) +
#   geom_point(aes(x = species, y = mean, color = metric)) +
#   geom_segment(aes(x = species, xend = species, 
#                    y = mean - se*1.96, yend = mean + se*1.6,
#                    color = metric)) +
#   coord_cartesian(ylim = c(0, 100), clip = "on") +
#   scale_color_manual(#labels = c("% Improvement in RMSPE \nfrom Conditioning",
#     #          "Fraction of Predictions\nImproved by Conditioning"),
#     values = c("deepskyblue", "coral3"), guide = "none") +
#   theme_bw() +
#   theme(strip.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.spacing = unit(1.5, "lines"),
#         plot.margin = margin(1,1,.5,.5, "cm")) +
#   labs(x = "species", y = "", color = "") + theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) + theme(
#     axis.text.x = element_blank(),
#     axis.title.x = element_blank(),
#     axis.ticks.x = element_blank()
#   )+ 
#   ggtitle('In-sample')+
#   theme(plot.title = element_text(hjust = 0.5))+theme(plot.margin=unit(c(0,0,0,0),"mm"))+ theme(plot.title = element_text(size=10))


imat <- as.data.frame(imat)
imat$species <- rownames(imat)
imat$`Frac u > c` <- imat$`Frac u > c`*100

imat1 <- imat[,c("Perc Diff", "Frac u > c", "species")]
colnames(imat1) <- c("percDiff", "percImpr", "species")
imat1 <- imat1 %>%
  pivot_longer(!species)
colnames(imat1) <- c("species", "metric", "oos")

colnames(all1) <- c("species", "metric", "is")
all1 <- all1[,1:3]


yzeros <- ydata
yzeros[yzeros > 0] <- 1
zeros <- as.data.frame(colMeans(yzeros))
zeros$species <- rownames(zeros)
colnames(zeros) <- c("abun", "species")

full <- full_join(all1, imat1, by = c("species", "metric")) %>%
  full_join(zeros, "species") %>%
  pivot_longer(cols = c("is", "oos"))


ggplot(full) + 
  geom_hline(yintercept = 50, color = "deepskyblue", alpha = 0.5) +
  geom_hline(yintercept = 0, color = "coral3", alpha = 0.5) +
  geom_point(aes(x = reorder(species, abun), y = value, color = metric, shape = name)) +
  scale_color_manual(#labels = c("% Improvement in RMSPE \nfrom Conditioning",
    #          "Fraction of Predictions\nImproved by Conditioning"),
    values = c("coral3","deepskyblue"), guide = "none") +
  scale_shape_manual(values = c(16, 1), guide = "none") +
  theme_bw() +
  theme(strip.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.spacing = unit(1.5, "lines"),
        plot.margin = margin(1,1,.5,.5, "cm"),
        axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  #coord_cartesian(ylim = c(0, 100), clip = "on") +
  labs(x = "", y = "")

ggsave("metrics.jpg", height = 5, width = 6, units = c("in"))

# imat1 <- imat %>% pivot_longer(cols = c("Unc MSPE", "Con MSPE", "Perc Diff", "Frac u > c"))
# 
# imat1$species <- tolower(imat1$species)
# imat1$species <- gsub("\\.", "\\ ", imat1$species)
# 
# imat2 <- 


# 
# 
# p2 <- imat1 %>% dplyr::filter(name == "Perc Diff" | name =="Frac u > c") %>%  
#   ggplot() +
#   geom_hline(yintercept = 50, color = "coral3", alpha = 0.5) +
#   geom_hline(yintercept = 0, color = "deepskyblue", alpha = 0.5) +
#   geom_point(aes(x = species, y = value, color = name))+
#   coord_cartesian(ylim = c(0, 100), clip = "on") +
#   scale_color_manual(#labels = c("% Improvement in RMSPE \nfrom Conditioning",
#     #          "Fraction of Predictions\nImproved by Conditioning"),
#     values = c("coral3","deepskyblue"), guide = "none") +
#   theme_bw() +
#   theme(strip.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.spacing = unit(1.5, "lines"),
#         plot.margin = margin(1,1,.5,.5, "cm")) +
#   labs(x = "species", y = "", color = "") + 
#   theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))+ 
#   ggtitle('Out-of-sample')+
#   theme(plot.title = element_text(hjust = 0.5))+
#   theme(plot.margin=unit(c(0,0,0,0),"mm"))+ 
#   theme(plot.title = element_text(size=10))
# 
# p2
# p1
# 
# p3 <- p1/p2
# p3
# 
# ggsave("in_sample_metrics.jpg", plot=p1, width = 15, height=10, units=c("cm"), dpi=500)
# 
# ggsave("out_of_sample_metrics.jpg", plot=p2, width = 15, height=10, units=c("cm"), dpi=500)
# 
# ggsave("all_metrics.jpg", plot=p3, width = 20, height=15, units=c("cm"), dpi=500)
# 


```

#4 Map
In sample map 
```{r}
## Map ----

load('out_full_CPUE_fall.Rdata')
output <- out1
ydata <- output$inputs$y
xdata <- output$inputs$xdata

total1 <- read.csv("total_with_env.csv") #this has the missing lat longs in it 

colnames(total1) <- gsub("\\_", ".", colnames(total1)) #gjam doesnt like _ in column names 
total1 <- total1[total1$year>1986,] 
total1 <- total1[total1$SEASON == "FALL",]

total1$SEDSIZE <- ifelse(total1$SEDIMENT ==  '0 - 0.03 Silt/Mud', .03, 
                        ifelse(total1$SEDIMENT ==  '0.03 - 0.17 Sand', .17,
                               ifelse(total1$SEDIMENT ==  '0.17 - 0.35 Sand',.35, 
                                      ifelse(total1$SEDIMENT ==  '0.35 - 0.36 Sand', .36,
                                             ifelse(total1$SEDIMENT ==  '0.36 - 0.48 Sand', .48, ifelse(total1$SEDIMENT ==  '0.48+ Coarse Sand to Gravel',.50,NA))))))

xnames <- c("depth", "BT", "BSAL", "SSAL", "SST", "chla", "month", "season", "SEDSIZE", "SUBREGION2", "haulid", "lat", "lon", "stratum", "stratumarea", "year")
xdata1 <- total1[colnames(total1) %in% xnames]
xdata1 <- xdata1[xdata1$SUBREGION2 == "Gulf of Maine",]
xdata1$SUBREGION2 <- gsub("Gulf of Maine", "GulfofMaine", xdata1$SUBREGION2 )

xdata1$month <- as.factor(xdata1$month)
xdata1$season <- as.factor(xdata1$season)
xdata1$SUBREGION2 <- as.factor(xdata1$SUBREGION2)


xdata <- left_join(xdata, xdata1,by = c("depth", "BT", "BSAL", "SSAL", "SST", "chla", "month", "season", "SEDSIZE", "SUBREGION2"))


#predicted biomass per stratum 
##observed
total_strat_obs <- cbind(xdata, ydata) 
total_strat_obs <- total_strat_obs %>% pivot_longer(cols = colnames(ydata), names_to = "spp", values_to = "biomass")

total_strat_obs$pred_typ <- "observed"

##conditional prediction
total_strat_con <- cbind(xdata, pred_con) 
total_strat_con <- total_strat_con %>% pivot_longer(cols = colnames(pred_con), names_to = "spp", values_to = "biomass")


total_strat_con$pred_typ <- "conditional"

##unconditional prediction
total_strat_uncon <- cbind(xdata, y_pred) 
total_strat_uncon <- total_strat_uncon %>% pivot_longer(cols = colnames(y_pred), names_to = "spp", values_to = "biomass")


total_strat_uncon$pred_typ <- "unconditional"

##diff 


total_strat <- rbind(total_strat_obs, total_strat_con, total_strat_uncon)



total_strat <- total_strat %>% group_by(season, spp, year, stratum, stratumarea, pred_typ) %>% summarise(biomass = mean(biomass)) %>% mutate(biomass_area = biomass*stratumarea)

total_strat <- total_strat %>% mutate(pred_type = factor(pred_typ)) %>% mutate(pred_type = fct_relevel(pred_type, c("observed", "unconditional", "conditional")))

#plotit 
strat <- sf::st_read("stratum/BTS_strata.shp")
#plot(strat)


strat <- left_join(strat, total_strat, by = c("STRATA" = "stratum"))
st_crs(strat) = 4326 #this is WGS 1984


# plot
library(rnaturalearth)
world <- ne_countries(scale = "medium", returnclass = "sf")
cols <- c("blue3", 
          "purple",
          "orange", "brown")
p <- ggplot() + geom_sf(data = world) + 
  geom_sf(data = subset(strat, strat$spp == "ATLANTIC.COD" & strat$year == 2016), aes(group = as.factor(biomass_area), fill = biomass_area))+ 
  coord_sf(xlim=c(-74, -65), ylim=c(40,46), expand = TRUE) + scale_fill_gradientn(colours = cols,
                        values = c(0, 0.1, 0.35, 0.75, 1)) + facet_wrap(~ pred_type) + theme_Publication() + theme(legend.position = "right", legend.direction = "vertical") +labs(fill = "Atlantic cod\nbiomass per stratum\n(kg/km2)")+
  theme_bw() +
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(hjust = 0),
        axis.title = element_blank(),
        panel.spacing.x = unit(6, "mm")) 

ggsave("codmaps_1.jpg", plot=p, width = 25, height=10, units=c("cm"), dpi=500)

#difference maps 
strat_sub <- strat %>% dplyr::select(-pred_typ, -biomass) %>% pivot_wider(names_from = "pred_type", values_from = biomass_area) %>% mutate(sub_con = conditional - observed, sub_uncon = unconditional - observed) %>% pivot_longer(cols = c("sub_uncon", "sub_con", "observed"))

strat_sub <- st_as_sf(strat_sub)
strat_sub <- strat_sub %>% filter(name %in% c("sub_uncon", "sub_con", "observed"))
st_crs(strat_sub) = 4326
strat_sub$name1 <- as.factor(strat_sub$name)

cols <- c("blue3", 
          "purple", "white",
          "orange", "brown")

label_names <- as_labeller(c("sub_uncon" = "unconditional - observed", "sub_con" = "conditional - observed", "observed" = "observed"))

p <- ggplot() + geom_sf(data = world) + 
  geom_sf(data = subset(strat_sub, strat_sub$spp == "ATLANTIC.COD" & strat_sub$year == 2012), aes(group = as.factor(value), fill = value))+ 
  coord_sf(xlim=c(-74, -65), ylim=c(40,46), expand = TRUE) + facet_wrap(~ name1, labeller = label_names) + theme_Publication() + theme(legend.position = "right", legend.direction = "vertical") +labs(fill = "Atlantic cod\nbiomass per stratum\n(kg/km2)")+
  theme_bw() +
  theme(panel.grid = element_blank(),
        strip.background = element_blank(),
        strip.text = element_text(hjust = 0),
        axis.title = element_blank(),
        panel.spacing.x = unit(6, "mm")) + scale_fill_gradient2(low = "blue3", high = "brown", mid = "white", midpoint = 1, breaks = c(-500,0,1000,2000))

p

ggsave("codmaps.jpg", plot=p, width = 25, height=10, units=c("cm"), dpi=500)
```
